Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-02-06 Manuscript/manuscriptPSsuppl.tex

\noindent
{\it POPC:POPS (5:1) mixtures with additional monovalent ions.}
Systems containing total 132 (110:22) lipids
were simulated with the additional potassium and sodium countertion concentrations
of $\sim$450 mM and 890 mM using CHARMM36 force field at 298 K. The data, force field and
input files generated using parameters given by CHARMM-GUI \cite{lee16,jo18} are
available from Refs. \citenum{charmm36pops+83popcT298Kwith450mMK,charmm36pops+83popcT298Kwith890mMK,charmm36pops+83popcT298Kwith440mMNa,charmm36pops+83popcT298Kwith890mMNa}. \\

@@

\subsection{MacRog}
\noindent
    {\it POPC bilayers.} Simulation data available from Ref. \citenum{macrogPOPC298K}.
    \todo{For example this data set has pretty good description of simulation details in the
      reference. I am not sure how much we should repeat these in here.}

The POPC bilayer consisting of a total of 128 lipids was constructed from single lipid structure taken 
from Ref.~\cite{kulig15c}. The bilayer was hydrated by a total of 5120 water 
molecules, corresponding to 40 water molecules per lipid.

The MacRog lipid parameters \cite{kulig15b}, obtained from 
\cite{kulig15c}, were used, along with the TIP3P water model 
\cite{jorgensen83}. The simulation parameters were identical to those used for 
POPC/POPC mixtures with the MacRog force field, see below. The simulation was run for 200~ns, 
out of which \hl{??~ns: (this is not tabulated in the MS or the SI)} were used for analyses.

@@

The bilayers containing a total of 120 POPC and 24 POPS molecules distributed evenly among the 
two leaflets were set up using CHARMM-GUI \cite{lee16,jo18}.
The bilayers were solvated by 5760 water molecules, corresponding to 40 water molecules for lipid.
Additionally, 24 potassium ions were added to neutralize the charge of the POPS head groups.

These bilayers were simulated in the presence of counterions only, as well as together with different
concentrations of CaCl$_2$ or KCl. For the former, concentrations of 100~mM, 300~mM, 1~M, and 3~M 
were considered, which corresponded to the amounts of 10/20, 31/62, 104/208, and 311/622 
of Ca$^{2+}$/Cl$^-$ ions. For membranes with KCl, concentrations of 500~mM, 1~M, 2~M, and 3~M of 
were considered, corresponding to 52, 104, 208, or 311 pairs of K$^+$ and Cl$^-$ ions.

@@

In all simulations, TIP3P~\cite{jorgensen1983comparison} was used as the solvent. SETTLE~\cite{miyamoto92} algorithm was employed 
to constrain the bonds in TIP3P water molecules. Lipid parameters were obtained from 
Amber Lipid17 force field. For the ions, one set of simulations were run with Joung-Cheatham
ion parameters~\cite{joung2008determination} and one set of simulations were run with ff99SB~\cite{aqvist90} ion parameters. The data in Table~\ref{PSsystems} and Refs~\citenum{lipid17POPSjcions,lipid17POPSff99ions,lipid17DOPSjcions,lipid17DOPSff99ions} contain the details of the simulations and the raw trajectories, respectively.

@@

The initial structure of the bilayers with 120 POPC and 24 POPS lipids were obtained from CHARMM-GUI~\cite{lee16,jo18}. Bilayers were solvated with 5760 TIP3P water molecules~\cite{jorgensen1983comparison}. For the simulations of the lipids and ions, the Amber Lipid 17 and ff99SB force fields were used, respectively. The simulation steps are the same as the the simulations of DOPS and POPS bilayers described above, with the only difference that the total simulation time was 300~ns, of which last 200~ns was used for the analysis. The number of ions in each specific system is given in Table~\ref{mixedIONsystemsMONOVALENT}. The trajectory files can be found in Refs~\citenum{POPCpopsLIPID17withKCI,POPCpopsLIPID17withK,POPCpopsLIPID17withNaCI,POPCpopsLIPID17withNa}.\\

@@

One set of simulations of POPC:POPS (5:1) mixtures with additional calcium ions has been performed using Amber18 simulation package. Number of ions and water molecules for this set of simulations are shown in Table~\ref{mixedIONsystems}. The simulation details, minimization, heating, and equilibration runs are the same as described above for the other POPC:POPS (5:1) simulations with Amber Lipid17 forcefield. For all simulations, ff99SB~\cite{aqvist90} ion parameters and Amber lipid 17 force field were used. For each ion concentration, 100~ns equilibration runs were followed by 300~ns production runs with two independent trajectories, of which last 200~ns were used for the analysis. The trajectory files can be found in Ref~\citenum{POPCpopsLIPID17withCaCl} 
\todo{B. Kav and J. Melcr, please write simulation details} \\

>>>>

\noindent
{\it POPC:POPS (5:1) mixtures with additional monovalent ions.}
POPC:POPS (110:22) mixtures containing total 132 lipids with the additional
potassium or sodium ions corresponding concentrations of $\sim$450 mM and 890 mM
were generated with the CHARMM-GUI \cite{lee16,jo18}. Systems were simulated using
Gromacs 5 \cite{abraham2015gromacs} and CHARMM36 force field with the simulation
parameters given by the CHARMM-GUI at 298 K. For further details, simulation files
and data, see table \ref{mixedIONsystemsMONOVALENT} and
Refs. \citenum{charmm36pops+83popcT298Kwith450mMK,charmm36pops+83popcT298Kwith890mMK,charmm36pops+83popcT298Kwith440mMNa,charmm36pops+83popcT298Kwith890mMNa}. \\

@@

\subsection{MacRog}
\noindent
{\it POPC bilayers.} 
% \todo{For example this data set has pretty good description of simulation details in the
% reference. I am not sure how much we should repeat these in here.}
The POPC bilayer consisting of a total of 128 lipids was constructed from single lipid structure taken 
from Ref.~\citenum{kulig15c}. The bilayer was hydrated by a total of 5120 water 
molecules, corresponding to 40 water molecules per lipid.
The MacRog lipid parameters \cite{kulig15b}, obtained from Ref.~\citenum{kulig15c}, were used,
along with the TIP3P water model~\cite{jorgensen83}. The simulation parameters were identical to those used for 
POPC/POPC mixtures with the MacRog force field, see below. The simulation was run for 200~ns using Gromacs 2016.3~\cite{abraham15}, 
out of which 150~ns were used for analyses. The simulation data and related files are available from Ref. \citenum{macrogPOPC298K}.

@@

The bilayers containing a total of 120 POPC and 24 POPS molecules distributed evenly among the 
two leaflets were set up using CHARMM-GUI \cite{lee16,jo18}.
The bilayers were solvated by 5760 water molecules, corresponding to 40 water molecules for lipid.
Additionally, 24 potassium ions were added to neutralize the charge of the POPS head groups.
These bilayers were simulated using Gromacs ?? \cite{check the version} in the presence of counterions only, as well as together with different
concentrations of CaCl$_2$ or KCl. For the former, concentrations of 100~mM, 300~mM, 1~M, and 3~M 
were considered, which corresponded to the amounts of 10/20, 31/62, 104/208, and 311/622 
of Ca$^{2+}$/Cl$^-$ ions. For membranes with KCl, concentrations of 500~mM, 1~M, 2~M, and 3~M of 
were considered, corresponding to 52, 104, 208, or 311 pairs of K$^+$ and Cl$^-$ ions.

@@

In all simulations, TIP3P~\cite{jorgensen1983comparison} was used as the solvent. SETTLE~\cite{miyamoto92} algorithm was employed 
to constrain the bonds in TIP3P water molecules. Simulations were ran using Amber18 simulation package \cite{??} \todo{Is this correct, do we need the citation?}
and lipid parameters were obtained from Amber Lipid17 force field \cite{gould18}. For the ions, one set of simulations were run with Joung-Cheatham
ion parameters~\cite{joung2008determination} and one set of simulations were run with ff99SB~\cite{aqvist90} ion parameters. The data in Table~\ref{PSsystems} and Refs~\citenum{lipid17POPSjcions,lipid17POPSff99ions,lipid17DOPSjcions,lipid17DOPSff99ions} contain the details of the simulations and the raw trajectories, respectively.

@@

The initial structure of the bilayers with 120 POPC and 24 POPS lipids were obtained from CHARMM-GUI~\cite{lee16,jo18}.
Bilayers were solvated with 5760 TIP3P water molecules~\cite{jorgensen1983comparison}. For the simulations of the lipids and ions,
the Amber Lipid 17 \cite{gould18} and ff99SB~\cite{aqvist90} force fields were used, respectively.
Simulations were ran using Amber18 simulation package \cite{??} \todo{Is this correct, do we need the citation?}
The simulation steps are the same as the the simulations of
DOPS and POPS bilayers described above, with the only difference that the total simulation time was 300~ns, of which last 200~ns
was used for the analysis. The number of ions in each specific system is given in Table~\ref{mixedIONsystemsMONOVALENT}.
The trajectory files can be found in Refs~\citenum{POPCpopsLIPID17withKCI,POPCpopsLIPID17withK,POPCpopsLIPID17withNaCI,POPCpopsLIPID17withNa}.\\

@@

One set of simulations of POPC:POPS (5:1) mixtures with additional calcium ions has been performed using Amber18 simulation package. Number of ions and water molecules for this set of simulations are shown in Table~\ref{mixedIONsystems}. The simulation details, minimization, heating, and equilibration runs are the same as described above for the other POPC:POPS (5:1) simulations with Amber Lipid17 forcefield. For all simulations, ff99SB~\cite{aqvist90} ion parameters and Amber lipid 17 force field were used. For each ion concentration, 100~ns equilibration runs were followed by 300~ns production runs with two independent trajectories, of which last 200~ns were used for the analysis. The trajectory files can be found in Ref~\citenum{POPCpopsLIPID17withCaCl}

\todo{J. Melcr, please write simulation details} \\
